---
title: "A case for systematic sound symbolism in pragmatics: Supporting information"
output:
    pdf_document:
        includes:
            in_header: header.tex 
        toc: true
---



```{r echo=F}
getMEText = function(r,ef, wald=NULL){

AIC = r[2,]$AIC
loglikDiff = signif(diff(r$logLik),2)
chi = round(r$Chisq[2],2)
df = r$`Chi Df`[2]
p = signif(r$`Pr(>Chisq)`[2],2)

wald.text = ""

if(!is.null(wald)){
  est = signif(wald[1],2)
  stder = signif(wald[2],2)
  t = signif(wald[3],2)
  wptext = ""
  if(!is.na(wald[4])){
    wptext = paste(", Wald p =",signif(wald[4],2))
  }
  wald.text = paste("beta = ",est,", std.err = ",stder, ", Wald t = ",t,wptext,';')
}

begin = 'There was no significant'
if(p <0.1){
  begin = "There was a marginal"
}
if(p < 0.05){
  begin = 'There was a significant'  
}


return(paste(begin,ef,"(",wald.text,"log likelihood difference =",
             loglikDiff,", df = ",df,", Chi Squared =", chi,", p = ",p,")."))
}

```

# Introduction

In this analysis, we leave out the first few trials.


# Load libraries

```{r warning=FALSE, message=FALSE}
library(lme4)
library(lattice)
library(gplots)
library(ggplot2)
library(sjPlot)
library(party)
library(Rmisc)
library(dplyr)
```

Function for converting from logit scale

```{r}
logit2per = function(X){
  return(exp(X)/(1+exp(X)))
}
```

```{r echo=FALSE}
# Set working directory
setwd("~/Documents/MPI/SemanticsPragmatics/2015Course/Projects/Slonimska/InitialPhonemeExperiment/SwitchboardSamplesExperiment/processing/")
```



# Load data

```{r}
d = read.csv("../Data/Lab_Processed.csv")
```

Make *answer* a binary variable.

```{r}
d$answer = d$answer=="Yes"
d$lastAnswer = d$lastAnswer=="Yes"
```

Relevel response phoneme and context.

```{r}
d$responsePhoneme = relevel(d$responsePhoneme, 'other')
d$context = relevel(d$context, 'ST')
```

Center trial number, so that the intercept will reflect probabilities in the middle of the experiment.

```{r}
d$trialNumber.center = d$trialNumber - 25
```



## Data exclusion

We exclude participant 13 because they took much longer than other participants.

```{r}
d = d[as.character(d$partID)!="13",]
```

\newpage



# Decision tree

In order to get an idea of the structure of the data, we make a binary decision tree based on the data.  We try to predict the participant's response by context, the first phoneme of the response, the type of turn the response was taken from and the participants' age, sex and the type of English they speak (American, British, or other, which are the main groups).

```{r}
cx = ctree(answer ~ 
             context +responsePhoneme + responseType +
             Age + Sex + EnglishType + 
             response.sex + context.sex +
             response.firstO + trialNumber + lastAnswer +
             blockName,
           data = d, 
           controls = ctree_control(mincriterion = 0.95))

```

Plot the decision tree:

```{r}
plot(cx, terminal_panel=node_barplot)
```

Context is the most important factor, followed by first phoneme of the response.

But we also see that there's a split by trial number.  In this analysis we leave out the first 5 trials for each participant:

```{r}
d = d[d$trialNumber >6,]
```


\newpage

# Mixed effects models

Make a series of mixed effects models.   We can fix this using the "bobyqa" optimiser for both phases of the convergence and letting the algorithm run longer:

```{r}
gcontrol = glmerControl(optimizer="bobyqa",optCtrl = list(maxfun=2e4))
```

(Note that several convergence algorithms were tested, and the three best fitting solutions had essentially no differences in fixed effect estimates)

## Random effects structure

We have a good idea of what the random effects structure should be, but first we check whether there are significant differences by participant etc.

```{r ranEfModels, cache=T}
mA0 = glmer(
  answer ~ 1 + 
    (1 | partID),   
  data = d,
  family = binomial,
  control = gcontrol
)


mA0b = glmer(
  answer ~ 1 + 
    (1 | blockName/partID) ,   
  data = d,
  family = binomial,
  control = gcontrol
)
anova(mA0,mA0b)
```

There is no significant improvement in the model when taking stimulus set into account. Because it complicates the analysis, we'll leave it out.

```{r ranefModels2, cache=T}
mA1 = glmer(
  answer ~ 1 + 
    (1 | partID) +   
    (1 | contextSample),  
  data = d,
  family = binomial,
  control = gcontrol
)

mA2 = glmer(
  answer ~ 1 + 
    (1 | partID) +   
    (1 | contextSample) +   
    (1 | responseSample),  
  data = d,
  family = binomial,
  control = gcontrol
)

mA3 = glmer(
  answer ~ 1 + 
    (1 + context| partID) +   
    (1 | contextSample) +   
    (1 | responseSample),   
  data = d,
  family = binomial,
  control = gcontrol
)

mA4 = glmer(
  answer ~ 1 + 
    (1 + context | partID) +
    (0 + responsePhoneme | partID) +
    (1 | contextSample) +  
    (1 | responseSample),  
  data = d,
  family = binomial,
  control = gcontrol
)

anova(mA0, mA1, mA2, mA3, mA4)

```

All proposed random effects significantly improve the fit of the model, except for the random slope for responsePhoneme by participant.  This variable nearly doubles the number of model parameters, so we will leave it out.

## Fixed effects

We are most interested in the effects of context and response type, but we need to check some other possible confounding variables.

*Trial*

```{r fixefModels,cache=T}
m0 = glmer(
  answer ~  1 + 
    (1 + context | partID) +   
    (0 + responsePhoneme | partID) +
    (1 | contextSample) +
    (1 | responseSample) ,   
  data = d,
  family = binomial,
  control = gcontrol
)

trial = glmer(
  answer ~ 1 + trialNumber.center +
    (1 + context | partID) +   
    (0 + responsePhoneme | partID) +
    (1 | contextSample) +
    (1 | responseSample),   
  data = d,
  family = binomial,
  control = gcontrol
)

trialQ = glmer(
  answer ~ 1 + trialNumber.center + I(trialNumber.center^2) +
    (1 + context | partID) +   
    (0 + responsePhoneme | partID) +
    (1 | contextSample) +
    (1 | responseSample),   
  data = d,
  family = binomial,
  control = gcontrol
)


anova(m0,trial, trialQ)
```

A significant effect of trial, but no significant quadratic term.

*Previous answer*

```{r fixefModels2, cache=T}
prevAns = glmer(
  answer ~ 1 + trialNumber.center + lastAnswer +
    (1 + context | partID) +   
    (0 + responsePhoneme | partID) +
    (1 | contextSample) +
    (1 | responseSample) ,   
  data = d,
  family = binomial,
  control = gcontrol
)
anova(trial,prevAns)

```

No significant effect of previous answer.

*Sex of speakers in samples*

```{r, fixefModels3,cache=T}

contS = glmer(
  answer ~ 1 + trialNumber.center + 
    context.sex +
    (1 + context | partID) +   
    (0 + responsePhoneme | partID) +
    (1 | contextSample) +
    (1 | responseSample) ,   
  data = d,
  family = binomial,
  control = gcontrol
)

respS = glmer(
  answer ~ 1 + trialNumber.center +  
    context.sex + response.sex +
    (1 + context | partID) +   
    (0 + responsePhoneme | partID) +
    (1 | contextSample) +
    (1 | responseSample),   
  data = d,
  family = binomial,
  control = gcontrol
)

contXrespS = glmer(
  answer ~ 1 + trialNumber.center +  
    context.sex * response.sex +
    (1 + context | partID) +   
    (0 + responsePhoneme | partID) +
    (1 | contextSample) +
    (1 | responseSample),   
  data = d,
  family = binomial,
  control = gcontrol
)


anova(trial,contS, respS, contXrespS)

```

No significant effects of the sex of the speakers in the samples.

*Sex of participants*

```{r fixefModels4, cache=T}
sex = glmer(
  answer ~ 1 + trialNumber.center + Sex +
    (1 + context | partID) +   
    (0 + responsePhoneme | partID) +
    (1 | contextSample) +
    (1 | responseSample) ,   
  data = d,
  family = binomial,
  control = gcontrol
)

anova(trial,sex)
```

No significant effect of the sex of the participant.

*Type of English spoken*

```{r fixefModels5, cache=T}
Etype = glmer(
  answer ~ 1 + trialNumber.center + EnglishType +
    (1 + context | partID) +   
    (0 + responsePhoneme | partID) +
    (1 | contextSample) +
    (1 | responseSample) ,   
  data = d,
  family = binomial,
  control = gcontrol
)

anova(trial,Etype)
```

No significant effec of the type of English the participant speaks.

### Effects of Context and Response

The only significant confounding variable is trial.

```{r fixefModels6, cache=T}

context = glmer(
  answer ~ 1  + trialNumber.center + 
    context +
    (1 + context | partID) +   
    (0 + responsePhoneme | partID) +
    (1 | contextSample) +
    (1 | responseSample) ,   
  data = d,
  family = binomial,
  control = gcontrol
)

rPhon = glmer(
  answer ~ 1 + trialNumber.center + 
    context + responsePhoneme +
    (1 + context | partID) +   
    (0 + responsePhoneme | partID) +
    (1 | contextSample) +
    (1 | responseSample),    
  data = d,
  family = binomial,
  control = gcontrol
)

conXrPh = glmer(
  answer ~ 1 + trialNumber.center +
    context * responsePhoneme +
    (1 + context | partID) +   
    (0 + responsePhoneme | partID) +
    (1 | contextSample) +
    (1 | responseSample),   
  data = d,
  family = binomial,
  control = gcontrol
)

anova(trial, context,rPhon, conXrPh)
```


*Interaction between Sex and responses*

```{r fixefModels7, cache=T}
Sex = glmer(
  answer ~ 1  + trialNumber.center +
    context * responsePhoneme +
    Sex +
    (1 + context | partID) +   
    (0 + responsePhoneme | partID) +
    (1 | contextSample) +
    (1 | responseSample),    
  data = d,
  family = binomial,
  control = gcontrol
)

SexXresp = glmer(
  answer ~ 1  + trialNumber.center +
    context * responsePhoneme +
    Sex*responsePhoneme +
    (1 + context | partID) +   
    (0 + responsePhoneme | partID) +
    (1 | contextSample) +
    (1 | responseSample) ,    
  data = d,
  family = binomial,
  control = gcontrol
)

SexXcon = glmer(
  answer ~ 1 + trialNumber.center +
    context * responsePhoneme +
    Sex*responsePhoneme +
    Sex:context +
    (1 + context | partID) +   
    (0 + responsePhoneme | partID) +
    (1 | contextSample) +
    (1 | responseSample) ,   
  data = d,
  family = binomial,
  control = gcontrol
)

SxXcoXre = glmer(
  answer ~ 1 + trialNumber.center +
    context * responsePhoneme +
    Sex*responsePhoneme*context +
    (1 + context | partID) +   
    (0 + responsePhoneme | partID) +
    (1 | contextSample) +
    (1 | responseSample) ,   
  data = d,
  family = binomial,
  control = gcontrol
)


```

```{r}
anova(conXrPh, Sex, SexXresp, SexXcon, SxXcoXre)
```

No effect by sex of participant.

\newpage

## Results

Model comparison

```{r}

anova(trial, context,rPhon, conXrPh)
```
\newpage

### Fixed effects

Model estimates:
  
```{r}
finalModel = conXrPh
save(finalModel, file="../results/FinalModel.Rdat")
summary(finalModel)
```

Relevel the response phoneme to see other comparisons:

```{r}
d2 = d
d2$responsePhoneme = relevel(d2$responsePhoneme,"wh")
fm2 = update(finalModel, data=d2)
summary(fm2)
```

Confidence intervals (through Wald method):

```{r}
CI = confint(finalModel,parm="beta_", method="Wald")
cx = summary(finalModel)$coef
cx = cbind(cx[,1],CI,cx[,2:4])
cx2 = cx
for(i in 1:5){cx2[,i] = round(cx2[,i],3)}
cx2
write.csv(cx, "../results/FinalModelCoefficients.csv")
```


\newpage

### Random effects

```{r}
dotplot(ranef(finalModel))
```


\newpage

## Summary

Here is a summary of the main results:

`r getMEText(anova(m0,context), "main effect of context")`

`r getMEText(anova(context,rPhon), "main effect of phoneme")`

`r getMEText(anova(rPhon,conXrPh), "interaction between context and phoneme")`

Work out model esimates for probabilities in each condition:
  
```{r}
# prob of responding 'yes' when:
# Context = ST, other response
logit2per(fixef(finalModel)[1])[[1]]
# Context = ST, no response 
logit2per(fixef(finalModel)[1] + fixef(finalModel)["responsePhonemenone"]  )[[1]]
# Context = ST, wh
logit2per(fixef(finalModel)[1] + fixef(finalModel)["responsePhonemewh"]  )[[1]]
# Context = IN, other response
logit2per(fixef(finalModel)[1] + fixef(finalModel)["contextIN"])[[1]]
# Context = IN, no response
logit2per(fixef(finalModel)[1] + 
            fixef(finalModel)["contextIN"] +
            fixef(finalModel)["responsePhonemenone"])[[1]]
# Context = IN, wh
logit2per(fixef(finalModel)[1] + 
            fixef(finalModel)["contextIN"] +
            fixef(finalModel)["responsePhonemewh"])[[1]]


```

\newpage

# Plots

Fixed effects estimates:
  
```{r}

feLabels = matrix(c(
  "(Intercept)"             ,"Intercept"      , NA,           
  "trialNumber.center", "Trial",NA,
  "contextST", "Context = Statement", "context",
  "contextIN", "Context = Initiating", "context",
  "responsePhonemenone", "no response", 'rPhon',
  "responsePhonemewh", "wh response", 'rPhon',
  "contextIN:responsePhonemenone", "Context = In: no response", "conXrPh",
  "contextIN:responsePhonemewh", "Context = In: wh response", "conXrPh"
), ncol=3, byrow = T)

feLabels2 = as.vector(feLabels[match(names(fixef(finalModel)),feLabels[,1]),2])


```



```{r finalModelFixedEffects, cache=T}
sjp.glmer(finalModel,'fe' , 
          show.intercept = T,
          geom.colors = c(1,1),
          axis.title = "Odds of selecting question",
          y.offset = 0.2
)
```


