---
title: "A case for systematic sound symbolism in pragmatics: Supporting information"
output:
    pdf_document:
        includes:
            in_header: header.tex 
        toc: true
---



```{r echo=F}
getMEText = function(r,ef, wald=NULL){

AIC = r[2,]$AIC
loglikDiff = signif(diff(r$logLik),2)
chi = round(r$Chisq[2],2)
df = r$`Chi Df`[2]
p = signif(r$`Pr(>Chisq)`[2],2)

wald.text = ""

if(!is.null(wald)){
  est = signif(wald[1],2)
  stder = signif(wald[2],2)
  t = signif(wald[3],2)
  wptext = ""
  if(!is.na(wald[4])){
    wptext = paste(", Wald p =",signif(wald[4],2))
  }
  wald.text = paste("beta = ",est,", std.err = ",stder, ", Wald t = ",t,wptext,';')
}

begin = 'There was no significant'
if(p <0.1){
  begin = "There was a marginal"
}
if(p < 0.05){
  begin = 'There was a significant'  
}


return(paste(begin,ef,"(",wald.text,"log likelihood difference =",
             loglikDiff,", df = ",df,", Chi Squared =", chi,", p = ",p,")."))
}

```

# Introduction

This is an analysis of an experiment into whether people can predict if an upcoming turn is a question or a statement, based on the previous turn type and the first phoneme of the target turn.

Participants listened to a series of audio samples.  Each audio sample was made up of a *context* by speaker 1 (Statement or Inititating turn) and a *response* by speaker 2.  The response was either no audio, a single segment [w] or a single semgent other than [w].


# Load libraries

```{r warning=FALSE, message=FALSE}
library(lme4)
library(lattice)
library(gplots)
library(ggplot2)
library(sjPlot)
library(party)
library(Rmisc)
library(dplyr)
```

Function for converting from logit scale

```{r}
logit2per = function(X){
  return(exp(X)/(1+exp(X)))
}
```

```{r echo=FALSE}
# Set working directory
setwd("~/Documents/MPI/SemanticsPragmatics/2015Course/Projects/Slonimska/InitialPhonemeExperiment/SwitchboardSamplesExperiment/processing/")
```



# Load data

```{r}
d = read.csv("../Data/Lab_Processed.csv")
```

Each row in the data is a single response from a participant to a single sample.  The key variables are:

-  *partID*:  identifies participants
-  *contextSample*: The name of the audio sample used for the context.
-  *responseSample*: The name of the audio sample used for the response.
-  *responsePhoneme*: The first segment of the response.
-  *responseType*: Whether the first segment of the response came from a question or statement.
-  *answer*: The participant's response to "Is the next turn a question?"

Make *answer* a binary variable.

```{r}
d$answer = d$answer=="Yes"
d$lastAnswer = d$lastAnswer=="Yes"
```

Relevel response phoneme and context.

```{r}
d$responsePhoneme = relevel(d$responsePhoneme, 'wh')
d$context = relevel(d$context, 'ST')
```

Center trial number, so that the intercept will reflect probabilities in the middle of the experiment.

```{r}
d$trialNumber.center = d$trialNumber - 25
# Scale between -1 and 1
d$trialNumber.center = d$trialNumber.center / 
  max(d$trialNumber.center)
```



## Data exclusion

We exclude participant 13 because they took much longer than other participants.

```{r}
d = d[as.character(d$partID)!="13",]
```

Are there any samples that look like outliers?  Make a basic model:

```{r cache=T}
m3 = glmer( 
  answer ~ 1 + context + responsePhoneme +
    (1 | partID) +   
    (1 | contextSample) +
    (1 | responseSample),   
  data = d,
  family = binomial,
  control = glmerControl(optimizer="bobyqa", optCtrl = list(maxfun=2e4))
)
```

Then look at the random effects.

```{r}
dotplot(ranef(m3))[[2]]
```

The sample "IN 18" is an outlier.  However, models have convergence problems when leaving it out.

The data has `r nrow(d)` observations:

```{r}
# Number of observations per participant
table(d$partID)
table(d$context, d$responsePhoneme )
```

\newpage

# Effects of block and trial

```{r}
plotmeans(answer ~ cut(trialNumber,seq(0,50,length.out = 11), include.lowest = T),
          ylab = "Prob of answering 'Question'",
          xlab = 'Trial',
          data = d[d$context=="ST",],ylim=c(0,1),
          col = 1, barcol = 1)
plotmeans(answer ~ cut(trialNumber,seq(0,50,length.out = 11), include.lowest = T),
          ylab = "Prob of answering 'Question'",
          xlab = 'Trial',
          data = d[d$context=="IN",],ylim=c(0,1),
          col = 2, barcol = 2, add=T)

plotmeans(d$answer ~ d$blockName,
          ylab = "Prob of answering 'Question'",
          xlab = 'Stimulus set',
          connect=F,
          ylim=c(0,1))

plotmeans(answer ~ lastAnswer,
          ylab = "Prob of answering 'Question'",
          xlab = "Previous response",
          legends = c("Not Q", "Question"),
          data = d)

```


# Decision tree

In order to get an idea of the structure of the data, we make a binary decision tree based on the data.  We try to predict the participant's response by context and the type of turn the response was taken from.

```{r}
cx.simple = ctree(answer ~ 
             context +responsePhoneme + responseType, data = d)
plot(cx.simple, terminal_panel=node_barplot(cx.simple))
```

And here is a more detailed analysis:

```{r}
cx = ctree(answer ~ 
             context +responsePhoneme + responseType +
             Age + Sex + EnglishType + 
             response.sex + context.sex +
             trialNumber + lastAnswer +
             blockName,
           data = d, 
           controls = ctree_control(mincriterion = 0.95))

```

Plot the decision tree:

```{r}
plot(cx, terminal_panel=node_barplot(cx, id=F))
```

Context is the most important factor, followed by first phoneme of the response.


\newpage

# Mixed effects models

Make a series of mixed effects models.   We can fix this using the "bobyqa" optimiser for both phases of the convergence and letting the algorithm run longer:

```{r}
gcontrol = glmerControl(optimizer="bobyqa",optCtrl = list(maxfun=2e4))
```

(Note that several convergence algorithms were tested, and the three best fitting solutions had essentially no differences in fixed effect estimates)

## Random effects structure

We have a good idea of what the random effects structure should be, but first we check whether there are significant differences by participant etc.

```{r ranEfModels, cache=T}
mA0 = glmer(
  answer ~ 1 +  
    (1 | partID),   
  data = d,
  family = binomial,
  control = gcontrol
)


mA0b = glmer(
  answer ~ 1 + 
    (1 | blockName/partID) ,   
  data = d,
  family = binomial,
  control = gcontrol
)
ltrf = anova(mA0,mA0b)
ltrf
```

There is no significant improvement in the model when taking stimulus set into account. Because it complicates the analysis, we'll leave it out.

```{r echo=F}
write.csv(as.data.frame(ltrf),"../results/lmerTests/RandomEffects.csv")
cat("Mixed effect models summary\n\n", 
    "../results/lmerTests/lmerTestSummary.txt")
rx  = getMEText(ltrf,"improvement in the fit of the model when adding a random effect for stimulus set")
cat(paste(rx,"\n\n",collapse=''), file="../results/lmerTests/lmerTestSummary.txt", append = T)
```



```{r ranefModels2, cache=T}
mA1 = glmer(
  answer ~ 1 +  
    (1 | partID) +   
    (1 | contextSample),  
  data = d,
  family = binomial,
  control = gcontrol
)

mA2 = glmer(
  answer ~ 1 + 
    (1 | partID) +   
    (1 | contextSample) +   
    (1 | responseSample),  
  data = d,
  family = binomial,
  control = gcontrol
)

mA3 = glmer(
  answer ~ 1 + 
    (1 + context| partID) +   
    (1 | contextSample) +   
    (1 | responseSample),   
  data = d,
  family = binomial,
  control = gcontrol
)

mA4 = glmer(
  answer ~ 1 + 
    (1 + context | partID) +
    (0 + responsePhoneme | partID) +
    (1 | contextSample) +  
    (1 | responseSample),  
  data = d,
  family = binomial,
  control = gcontrol
)

ltrf2 = anova(mA0, mA1, mA2, mA3, mA4)
ltrf2
```

All proposed random effects significantly improve the fit of the model, except for the random slope for responsePhoneme by participant.  Adding the correlation term between random effects slopes causes serious convergence issues, so we leave that out.

```{r echo=F}
write.csv(as.data.frame(ltrf2),"../results/lmerTests/RandomEffects2.csv")
rx  = getMEText(anova(mA0,mA1),"improvement in the fit of the model when adding a random effect for context sample")
cat(paste(rx,"\n\n",collapse=''), file="../results/lmerTests/lmerTestSummary.txt", append = T)
rx  = getMEText(anova(mA1,mA2),"improvement in the fit of the model when adding a random effect for response sample")
cat(paste(rx,"\n\n",collapse=''), file="../results/lmerTests/lmerTestSummary.txt", append = T)
rx  = getMEText(anova(mA2,mA3),"improvement in the fit of the model when adding a random slope for context by participant")
cat(paste(rx,"\n\n",collapse=''), file="../results/lmerTests/lmerTestSummary.txt", append = T)
rx  = getMEText(anova(mA2,mA3),"improvement in the fit of the model when adding a random slope for response phoneme by participant")
cat(paste(rx,"\n\n",collapse=''), file="../results/lmerTests/lmerTestSummary.txt", append = T)

```


## Fixed effects

We are most interested in the effects of context and response type, but we need to check some other possible confounding variables.

*Trial*

```{r fixefModels,cache=T}
m0 = glmer(
  answer ~ 1 +  
    (1 + context | partID) +   
    (0 + responsePhoneme | partID) +
    (1 | contextSample) +
    (1 | responseSample) ,   
  data = d,
  family = binomial,
  control = gcontrol
)

trial = glmer(
  answer ~ 1 + trialNumber.center +
    (1 + context | partID) +   
    (0 + responsePhoneme | partID) +
    (1 | contextSample) +
    (1 | responseSample),   
  data = d,
  family = binomial,
  control = gcontrol
)

trialQ = glmer(
  answer ~ 1 + trialNumber.center + I(trialNumber.center^2) +
    (1 + context | partID) +   
    (0 + responsePhoneme | partID) +
    (1 | contextSample) +
    (1 | responseSample),   
  data = d,
  family = binomial,
  control = gcontrol
)


lttr = anova(m0,trial, trialQ)
lttr
```

A significant effect of trial, but no significant quadratic term.

```{r echo=F}
write.csv(as.data.frame(lttr),"../results/lmerTests/Trial.csv")
rx  = getMEText(anova(m0,trial),"effect of trial")
cat(paste(rx,"\n\n",collapse=''), file="../results/lmerTests/lmerTestSummary.txt", append = T)
rx  = getMEText(anova(trial,trialQ),"quadratic effect of trial")
cat(paste(rx,"\n\n",collapse=''), file="../results/lmerTests/lmerTestSummary.txt", append = T)
```

*Previous answer*

```{r fixefModels2, cache=T}
prevAns = glmer( 
  answer ~ 1 + trialNumber.center + lastAnswer +
    (1 + context | partID) +   
    (0 + responsePhoneme | partID) +
    (1 | contextSample) +
    (1 | responseSample) ,   
  data = d,
  family = binomial,
  control = gcontrol
)
ltpa = anova(trial,prevAns)
ltpa
```

No significant effect of previous answer.

```{r echo=F}
write.csv(as.data.frame(ltpa),"../results/lmerTests/PreviousAnswer.csv")
rx  = getMEText(ltpa,"effect of previous answer")
cat(paste(rx,"\n\n",collapse=''), file="../results/lmerTests/lmerTestSummary.txt", append = T)
```

*Sex of speakers in samples*

```{r, fixefModels3,cache=T}

contS = glmer( 
  answer ~ 1 + trialNumber.center + 
    context.sex +
    (1 + context | partID) +   
    (0 + responsePhoneme | partID) +
    (1 | contextSample) +
    (1 | responseSample) ,   
  data = d,
  family = binomial,
  control = gcontrol
)

respS = glmer(
  answer ~ 1 + trialNumber.center +  
    context.sex + response.sex +
    (1 + context | partID) +   
    (0 + responsePhoneme | partID) +
    (1 | contextSample) +
    (1 | responseSample),   
  data = d,
  family = binomial,
  control = gcontrol
)

contXrespS = glmer(
  answer ~ 1 + trialNumber.center +  
    context.sex * response.sex +
    (1 + context | partID) +   
    (0 + responsePhoneme | partID) +
    (1 | contextSample) +
    (1 | responseSample),   
  data = d,
  family = binomial,
  control = gcontrol
)


ltsx = anova(trial,contS, respS, contXrespS)
ltsx

```

No significant effects of the sex of the speakers in the samples.

```{r echo=F}
write.csv(as.data.frame(ltsx),"../results/lmerTests/SexOfSpeakers.csv")
rx  = getMEText(anova(trial,contS),"effect of context speaker sex")
cat(paste(rx,"\n\n",collapse=''), file="../results/lmerTests/lmerTestSummary.txt", append = T)
rx  = getMEText(anova(contS,respS),"response speaker sex")
cat(paste(rx,"\n\n",collapse=''), file="../results/lmerTests/lmerTestSummary.txt", append = T)
rx  = getMEText(anova(respS, contXrespS),"interaction between context speaker sex and response speaker sex")
cat(paste(rx,"\n\n",collapse=''), file="../results/lmerTests/lmerTestSummary.txt", append = T)
```

*Sex of participants*

```{r fixefModels4, cache=T}
sex = glmer( 
  answer ~ 1 + trialNumber.center + Sex +
    (1 + context | partID) +   
    (0 + responsePhoneme | partID) +
    (1 | contextSample) +
    (1 | responseSample) ,   
  data = d,
  family = binomial,
  control = gcontrol
)

ltsxp = anova(trial,sex)
ltsxp
```

No significant effect of the sex of the participant.

```{r echo=F}
write.csv(as.data.frame(ltsxp),"../results/lmerTests/SexOfParticipants.csv")
rx  = getMEText(ltsxp,"effect of participant sex")
cat(paste(rx,"\n\n",collapse=''), file="../results/lmerTests/lmerTestSummary.txt", append = T)
```


*Age of participants*

```{r}
age = glmer( 
  answer ~ 1 + trialNumber.center + Age +
    (1 + context | partID) +   
    (0 + responsePhoneme | partID) +
    (1 | contextSample) +
    (1 | responseSample) ,   
  data = d,
  family = binomial,
  control = gcontrol
)

ltag = anova(trial,age)
ltag
```

No significant effect of age of partcipant.

```{r echo=F}
write.csv(as.data.frame(ltag),"../results/lmerTests/AgeOfParticipants.csv")
rx  = getMEText(ltag,"effect of participant age")
cat(paste(rx,"\n\n",collapse=''), file="../results/lmerTests/lmerTestSummary.txt", append = T)
```

*Type of English spoken*

```{r fixefModels5, cache=T}
Etype = glmer( 
  answer ~ 1 + trialNumber.center + EnglishType +
    (1 + context | partID) +   
    (0 + responsePhoneme | partID) +
    (1 | contextSample) +
    (1 | responseSample) ,   
  data = d,
  family = binomial,
  control = gcontrol
)

lten = anova(trial,Etype)
lten
```

No significant effec of the type of English the participant speaks.

```{r echo=F}
write.csv(as.data.frame(lten),"../results/lmerTests/DialectOfParticipants.csv")
rx  = getMEText(lten,"effect of dialect")
cat(paste(rx,"\n\n",collapse=''), file="../results/lmerTests/lmerTestSummary.txt", append = T)
```

### Effects of Context and Response

The only significant confounding variable is trial, so that forms the baseline.

```{r fixefModels6, cache=T}

context = glmer( 
  answer ~ 1 + trialNumber.center + 
    context +
    (1 + context | partID) +   
    (0 + responsePhoneme | partID) +
    (1 | contextSample) +
    (1 | responseSample) ,   
  data = d,
  family = binomial,
  control = gcontrol
)

rPhon = glmer(
  answer ~ 1 + trialNumber.center + 
    context + responsePhoneme +
    (1 + context | partID) +   
    (0 + responsePhoneme | partID) +
    (1 | contextSample) +
    (1 | responseSample),    
  data = d,
  family = binomial,
  control = gcontrol
)

conXrPh = glmer(
  answer ~ 1 + trialNumber.center +
    context * responsePhoneme +
    (1 + context | partID) +   
    (0 + responsePhoneme | partID) +
    (1 | contextSample) +
    (1 | responseSample),   
  data = d,
  family = binomial,
  control = gcontrol
)

anova(trial, context,rPhon, conXrPh)
```

*Interaction between Sex and responses*

```{r fixefModels7, cache=T}
Sex = glmer( 
  answer ~ 1 + trialNumber.center +
    context * responsePhoneme +
    Sex +
    (1 + context | partID) +   
    (0 + responsePhoneme | partID) +
    (1 | contextSample) +
    (1 | responseSample),    
  data = d,
  family = binomial,
  control = gcontrol
)

SexXresp = glmer(
  answer ~ 1 + trialNumber.center +
    context * responsePhoneme +
    Sex*responsePhoneme +
    (1 + context | partID) +   
    (0 + responsePhoneme | partID) +
    (1 | contextSample) +
    (1 | responseSample) ,    
  data = d,
  family = binomial,
  control = gcontrol
)

SexXcon = glmer(
  answer ~ 1 + trialNumber.center +
    context * responsePhoneme +
    Sex*responsePhoneme +
    Sex:context +
    (1 + context | partID) +   
    (0 + responsePhoneme | partID) +
    (1 | contextSample) +
    (1 | responseSample) ,   
  data = d,
  family = binomial,
  control = gcontrol
)

SxXcoXre = glmer(
  answer ~ 1 + trialNumber.center +
    context * responsePhoneme +
    Sex*responsePhoneme*context +
    (1 + context | partID) +   
    (0 + responsePhoneme | partID) +
    (1 | contextSample) +
    (1 | responseSample) ,   
  data = d,
  family = binomial,
  control = gcontrol
)


```

```{r}
ltsxx = anova(conXrPh, Sex, SexXresp, SexXcon, SxXcoXre)
ltsxx
```

No effect by sex of participant.

```{r echo=F}
write.csv(as.data.frame(ltsxx),"../results/lmerTests/SexInteractions.csv")
rx  = getMEText(anova(SexXresp , Sex),"interaction between sex of participant and response phoneme")
cat(paste(rx,"\n\n",collapse=''), file="../results/lmerTests/lmerTestSummary.txt", append = T)
rx  = getMEText(anova(SexXresp , SexXcon),"interaction between sex of participant and context")
cat(paste(rx,"\n\n",collapse=''), file="../results/lmerTests/lmerTestSummary.txt", append = T)
rx  = getMEText(anova(SxXcoXre , SexXcon),"three way interaction between sex of participant, context and response phoneme")
cat(paste(rx,"\n\n",collapse=''), file="../results/lmerTests/lmerTestSummary.txt", append = T)
```

*Interaction with trial*

```{r trialInteraction, cache=T}
trialXCon = glmer( 
  answer ~ 1 + trialNumber.center +
    context * responsePhoneme +
    trialNumber.center:context +
    (1 + context | partID) +   
    (0 + responsePhoneme | partID) +
    (1 | contextSample) +
    (1 | responseSample),   
  data = d,
  family = binomial,
  control = gcontrol
)

trialXph = glmer(
  answer ~ 1 + trialNumber.center +
    context * responsePhoneme +
    trialNumber.center:context +
    trialNumber.center:responsePhoneme +
    (1 + context | partID) +   
    (0 + responsePhoneme | partID) +
    (1 | contextSample) +
    (1 | responseSample),   
  data = d,
  family = binomial,
  control = gcontrol
)

trXcoXph = glmer(
  answer ~ 1 + trialNumber.center *
    context * responsePhoneme +
    (1 + context | partID) +   
    (0 + responsePhoneme | partID) +
    (1 | contextSample) +
    (1 | responseSample),   
  data = d,
  family = binomial,
  control = gcontrol
)

```

```{r trialInteractionAnova, cache=T}
lttrx = anova(conXrPh, trialXCon, trialXph, trXcoXph)
lttrx
```

Significant interaction between trial and context, but no reliable further interaction.

```{r echo=F}
write.csv(as.data.frame(lttrx),"../results/lmerTests/TrialInteractions.csv")
rx  = getMEText(anova(conXrPh, trialXCon),"interaction between trial and context")
cat(paste(rx,"\n\n",collapse=''), file="../results/lmerTests/lmerTestSummary.txt", append = T)
rx  = getMEText(anova(trialXCon, trialXph),"interaction between trial and response phoneme")
cat(paste(rx,"\n\n",collapse=''), file="../results/lmerTests/lmerTestSummary.txt", append = T)
rx  = getMEText(anova(trialXph, trXcoXph),"three way interaction between trial, context and response phoneme")
cat(paste(rx,"\n\n",collapse=''), file="../results/lmerTests/lmerTestSummary.txt", append = T)
```

\newpage

*Effect of stimulus set*

Below we adjust the random effects so that participants are nested within stimulus set (the variable *blockName*).  

```{r stimSetTest, cache=T}
stimSet = glmer(
     answer ~ 1 + trialNumber.center +
         context * responsePhoneme +
         trialNumber.center:context +
         (1 + context | blockName/partID) +   
         (0 + responsePhoneme | blockName/partID) +
         (1 | contextSample) +
         (1 | responseSample),   
     data = d,
     family = binomial,
     control = gcontrol
 )
```

```{r stimSetAnova, cache=T}
anova(trialXCon,stimSet)
```

There is no significant improvement in the model, and in any case the stimuli sets are counterbalanced experimentally, so we don't include it.  

In any case, the qualitative results are the same, and the estimates are very similar, suggesting that stimulus set does not have an impact on the main findings.

```{r}
cbind(without=fixef(trialXCon),withRForStimSet=fixef(stimSet))
```

\newpage

## Check /h/ phoneme samples

Only one stimuli set had a /h/ response phoneme, so we re-run the main analysis without those trials.

```{r checkH, cache=T}

takeOutSet = d[d$response.firstO=='h',]$setNum[1]

trialH = glmer(
  answer ~ 1 + trialNumber.center +
    (1 + context | partID) +   
    (0 + responsePhoneme | partID) +
    (1 | contextSample) +
    (1 | responseSample),   
  data = d[d$setNum != takeOutSet,],
  family = binomial,
  control = gcontrol
)

contextH = glmer( 
  answer ~ 1 + trialNumber.center + 
    context +
    (1 + context | partID) +   
    (0 + responsePhoneme | partID) +
    (1 | contextSample) +
    (1 | responseSample) ,   
  data = d[d$setNum != takeOutSet,],
  family = binomial,
  control = gcontrol
)

rPhonH = glmer(
  answer ~ 1 + trialNumber.center + 
    context + responsePhoneme +
    (1 + context | partID) +   
    (0 + responsePhoneme | partID) +
    (1 | contextSample) +
    (1 | responseSample),    
  data = d[d$setNum != takeOutSet,],
  family = binomial,
  control = gcontrol
)

conXrPhH = glmer(
  answer ~ 1 + trialNumber.center +
    context * responsePhoneme +
    (1 + context | partID) +   
    (0 + responsePhoneme | partID) +
    (1 | contextSample) +
    (1 | responseSample),   
  data = d[d$setNum != takeOutSet,],
  family = binomial,
  control = gcontrol
)

trialXConH = glmer( 
  answer ~ 1 + trialNumber.center +
    context * responsePhoneme +
    trialNumber.center:context +
    (1 + context | partID) +   
    (0 + responsePhoneme | partID) +
    (1 | contextSample) +
    (1 | responseSample),   
  data = d[d$setNum != takeOutSet,],
  family = binomial,
  control = gcontrol
)

anova(trialH, contextH,rPhonH, conXrPhH, trialXConH)

summary(trialXConH)$coef
```

There are no qualitative differences when removing these trials.

\newpage

## Results

Model comparison

```{r}
mainResults = anova(m0, trial, context,rPhon, conXrPh, trialXCon)
mainResults
```

```{r echo=F}
write.csv(as.data.frame(mainResults), "../results/MainAnovaTest.csv")
```


\newpage

### Fixed effects

Model estimates:
  
```{r}
finalModel = trialXCon
save(finalModel, file="../results/FinalModel.Rdat")
summary(finalModel)
```

Relevel the response phoneme to see other comparisons:

```{r}
d2 = d
d2$responsePhoneme = relevel(d2$responsePhoneme,"other")
fm2 = update(finalModel, data=d2)
summary(fm2)

write.csv(as.data.frame(summary(fm2)$coef),
          "../results/FinalModelCoefficients_relevel.csv")
```

Confidence intervals (through Wald method):

```{r}
CI = confint(finalModel,parm="beta_", method="Wald")
cx = summary(finalModel)$coef
cx = cbind(cx[,1],CI,cx[,2:4])
cx2 = cx
for(i in 1:5){cx2[,i] = round(cx2[,i],3)}
cx2 = as.data.frame(cx2)
names(cx2)[1] = "estimate.logit"
cx2$esimate.odds = exp(cx2[,1])
cx2$esimate.odds.lower = exp(cx2[,2])
cx2$esimate.odds.upper = exp(cx2[,2])

cx2

write.csv(cx, "../results/FinalModelCoefficients.csv")
```


\newpage

### Random effects

```{r}
dotplot(ranef(finalModel))
x = as.data.frame(ranef(finalModel)$partID)
rownames(x) = rownames(ranef(finalModel)$partID)
for(i in 1:ncol(x)){
  x = x[order(x[,i]),]
  plot(x[,i],1:nrow(x),
       main = paste("PartID,",colnames(x)[i]),
       yaxt='n',
       ylab='partID', xlab='')
  abline(v=0)
  axis(2,at=1:nrow(x),labels=rownames(x), las=2)
}


```

Note that the intercept for partID is exactly correlated with the random slope for contextIN by partID.

\newpage

## Summary

Here is a summary of the main results:

`r getMEText(anova(trial,context), "main effect of context")`

`r getMEText(anova(context,rPhon), "main effect of phoneme")`

`r getMEText(anova(rPhon,conXrPh), "interaction between context and phoneme")`

`r getMEText(anova(m0, trial), "main effect of trial")`



Work out model esimates for probabilities in each condition:
  
```{r}
# prob of responding 'yes' when:
# Context = ST, other response
logit2per(fixef(finalModel)[1])[[1]]
# Context = ST, no response 
logit2per(fixef(finalModel)[1] + fixef(finalModel)["responsePhonemenone"]  )[[1]]
# Context = ST, wh
logit2per(fixef(finalModel)[1] + fixef(finalModel)["responsePhonemewh"]  )[[1]]
# Context = IN, other response
logit2per(fixef(finalModel)[1] + fixef(finalModel)["contextIN"])[[1]]
# Context = IN, no response
logit2per(fixef(finalModel)[1] + 
            fixef(finalModel)["contextIN"] +
            fixef(finalModel)["responsePhonemenone"])[[1]]
# Context = IN, wh
logit2per(fixef(finalModel)[1] + 
            fixef(finalModel)["contextIN"] +
            fixef(finalModel)["responsePhonemewh"])[[1]]


```

\newpage

# Plots

Fixed effects estimates:
  
```{r}

feLabels = matrix(c(
  "(Intercept)"             ,"Intercept"      , NA,           
  "trialNumber.center", "Trial",NA,
  "contextST", "Context = Statement", "context",
  "contextIN", "Context = Initiating", "context",
  "responsePhonemenone", "No response", 'rPhon',
  "responsePhonemewh", "wh response", 'rPhon',
  "responsePhonemeother","Other response", 'rPhon',
  "contextIN:responsePhonemenone", "Context = In: no response", "conXrPh",
  "contextIN:responsePhonemewh", "Context = In: wh response", "conXrPh",
  "contextIN:responsePhonemeother", "Context = In: other response", "conXrPh",
  "trialNumber.center:contextIN","Trial:Context = In",'trialXCon'
), ncol=3, byrow = T)

feLabels2 = as.vector(feLabels[match(names(fixef(finalModel)),feLabels[,1]),2])


```



```{r finalModelFixedEffects, cache=T}
sjp.glmer(finalModel,'fe',  
          show.intercept = T,
          geom.colors = c(1,1),
          axis.title = "Odds of selecting question",
          y.offset = 0.2,
          axis.labels = feLabels2[2:length(feLabels2)]
)
```



\newpage

## Raw data plots

```{r rawDataPlots, cache=T}

d$responsePhoneme2 = relevel(relevel(d$responsePhoneme,'other'),'none')

sumStats = group_by(d, partID ,context,responsePhoneme2 ) %>%
              summarise(mean =mean(answer) )

sumStats2 = summarySE(sumStats, measurevar="mean", groupvars=c("context","responsePhoneme2"))
sumStats2$upper = sumStats2$mean + sumStats2$ci
sumStats2$lower = sumStats2$mean - sumStats2$ci

sumStats2

dodge <- position_dodge(width=0.5) 

main.plot <- ggplot(sumStats2,
	aes(x = responsePhoneme2, y = mean, colour=context)) +
  geom_point() + geom_line(aes(group=context)) +
  geom_errorbar(aes(ymax=mean+ci, ymin=mean-ci), width=0.25) +
  xlab("First Phoneme in Response") +
  ylab("Proportion of 'Question' answers") +
  coord_cartesian(ylim=c(0,1)) + 
  scale_color_discrete(breaks=c("ST","IN"),
                       labels=c("Statement","Initiating"),
                       name="Context")

main.plot

pdf("../results/graphs/PropQResponses_by_firstPhoneme_withPartCI.pdf",
    width = 4, height=3)
main.plot
dev.off()


main.plot2 <- ggplot(sumStats2,
	aes(x = context, y = mean, colour=responsePhoneme2)) +
  geom_point(position=dodge) + geom_line(aes(group=responsePhoneme2), position=dodge) +
  geom_errorbar(aes(ymax=mean+ci, ymin=mean-ci), width=0.25, position=dodge) +
  xlab("Previous turn type (context)") +
  ylab("Proportion of 'Question' answers") +
  coord_cartesian(ylim=c(0,1)) + 
  scale_color_discrete(breaks=c("none","other",'wh'),
                       labels=c("None","Non-wh","wh"),
                       name="First phoneme") +
  scale_x_discrete(breaks=c("ST", "IN"),
                      labels=c("Statement", "Initiating"))

main.plot2

pdf("../results/graphs/PropQResponses_by_context_withPartCI.pdf",
    width = 4, height=3)
main.plot2
dev.off()

```

\newpage

# Predicting response type


```{r}
d2 = d[d$responsePhoneme!="none",]

table(d2$answer,d2$responseType)

d2$correct = "Correct"
d2$correct[!d2$answer & d2$responseType=="Q"] = "Incorrect"
d2$correct[d2$answer & d2$responseType=="other"] = "Incorrect"
# number of "correct" responses
table(d2$correct)

```

```{r}
m0T = glmer(answer ~ 1 + context*responsePhoneme +
          (1 + context | partID) +   
          (1 | contextSample) +
          (1 | responseSample) ,
           data = d2,
           family = binomial,
          control=gcontrol)

respT = glmer(answer ~ 1 + context*responsePhoneme +
                responseType +
          (1 + context | partID) +   
          (1 | contextSample) +
          (1 | responseSample) ,
           data = d2,
           family = binomial,
          control=gcontrol)

respTXco = glmer(answer ~ 1 + context*responsePhoneme +
                responseType*context +
          (1 + context | partID) +   
          (1 | contextSample) +
          (1 | responseSample) ,
           data = d2,
           family = binomial,
          control=gcontrol)

respTXrp = glmer(answer ~ 1 + context*responsePhoneme +
                responseType*context +
                + responseType: responsePhoneme +
          (1 + context | partID) +   
          (1 | contextSample) +
          (1 | responseSample) ,
           data = d2,
           family = binomial,
          control=gcontrol)

rTXcoXrp = glmer(answer ~ 1 + context*responsePhoneme +
                responseType*context *responsePhoneme +
          (1 + context | partID) +   
          (1 | contextSample) +
          (1 | responseSample) ,
           data = d2,
           family = binomial,
          control=gcontrol)

anova(m0T, respT, respTXco, respTXrp, rTXcoXrp)

```

No effects of actual response type.



